Evaluating the Decency and Consistency of Data Validation Tests Generated by LLMs

### 2023-10-01 paper/paper.qmd

---
title: "Evaluating the Decency and Consistency of Data Validation Tests Generated by LLMs"
author: 
    - name: Rohan Alexander^[Information and Statistical Sciences, University of Toronto. rohan.alexander@utoronto.ca.]
    - name: Lindsay Katz^[The Investigative Journalism Foundation]
    - name: Callandra Moore^[The Investigative Journalism Foundation]
    - name: Zane Schwartz^[The Investigative Journalism Foundation]
date: today
date-format: long
abstract: "We investigated the potential of Large Language Models (LLMs) in developing dataset validation tests. We carried out 96 experiments each for both GPT-3.5 and GPT-4, examining different prompt scenarios, learning modes, temperature settings, and roles. The prompt scenarios were: 1) Asking for expectations, 2) Asking for expectations with a given context, 3) Asking for expectations after requesting a simulation, and 4) Asking for expectations with a provided data sample. For learning modes, we tested: 1) zero-shot, 2) one-shot, and 3) few-shot learning. We also tested four temperature settings: 0, 0.4, 0.6, and 1. Furthermore, two distinct roles were considered: 1) “helpful assistant”, 2) “expert data scientist”. To gauge consistency, every setup was tested five times. The LLM-generated responses were benchmarked against a gold standard suite, created by an experienced data scientist knowledgeable about the data in question. We find there are considerable returns to the use of few-shot learning, and that the more explicit the data setting can be the better. The results of the best combinations complement, but do not replace, those of the gold standard. The best LLM configurations complement, rather than substitute, the gold standard results. This study underscores the value LLMs can bring to the data cleaning and preparation stages of the data science workflow."
thanks: "We thank Monica Alexander. Code and data: https://github.com/RohanAlexander/evaluating_decency_and_consistency. Contributions: RA had the initial idea. RA and CM developed the experimental set-up. ZS established the political donations datasets. LK developed the initial suite of dataset validation tests that we compare the LLMs against. RA wrote the code to interact with the LLMs, evaluated the LLM outputs, and did the modelling. All authors contributed to writing the initial draft as well as improving and finalizing the paper."
bibliography: "bibliography.bib"
format: pdf
number-sections: true
---

>>>>

---
title: "Evaluating the Decency and Consistency of Data Validation Tests Generated by LLMs"
author: 
    - name: Rohan Alexander^[Information and Statistical Sciences, University of Toronto. rohan.alexander@utoronto.ca.]
    - name: Lindsay Katz^[The Investigative Journalism Foundation]
    - name: Callandra Moore^[The Investigative Journalism Foundation]
    - name: Zane Schwartz^[The Investigative Journalism Foundation]
date: today
date-format: long
abstract: "We investigated the potential of large language models (LLMs) in developing dataset validation tests. We carried out 96 experiments each for both GPT-3.5 and GPT-4, examining different prompt scenarios, learning modes, temperature settings, and roles. The prompt scenarios were: 1) Asking for expectations, 2) Asking for expectations with a given context, 3) Asking for expectations after requesting a simulation, and 4) Asking for expectations with a provided data sample. For learning modes, we tested: 1) zero-shot, 2) one-shot, and 3) few-shot learning. We also tested four temperature settings: 0, 0.4, 0.6, and 1. Furthermore, two distinct roles were considered: 1) “helpful assistant”, 2) “expert data scientist”. To gauge consistency, every setup was tested five times. The LLM-generated responses were benchmarked against a gold standard suite, created by an experienced data scientist knowledgeable about the data in question. We find there are considerable returns to the use of few-shot learning, and that the more explicit the data setting can be the better. The results of the best combinations complement, but do not replace, those of the gold standard. The best LLM configurations complement, rather than substitute, the gold standard results. This study underscores the value LLMs can bring to the data cleaning and preparation stages of the data science workflow."
thanks: "We thank Monica Alexander. Code and data: https://github.com/RohanAlexander/evaluating_decency_and_consistency. Contributions: RA had the initial idea. RA and CM developed the experimental set-up. ZS established the political donations datasets. LK developed the initial suite of dataset validation tests that we compare the LLMs against. RA wrote the code to interact with the LLMs, evaluated the LLM outputs, and did the modelling. All authors contributed to writing the initial draft as well as improving and finalizing the paper."
bibliography: "bibliography.bib"
format: pdf
number-sections: true
---
